%%
%% cat.tex
%% 
%% Made by Alex Nelson <pqnelson@gmail.com>
%% Login   <alex@lisp>
%% 
%% Started on  2025-08-05T09:47:03-0700
%% Last update 2025-08-05T09:47:03-0700
%% 

\chapter{Category Theory}
\needs{graph.tex}

\section{Basic Definitions}

\begin{remark}[Category is a cryptomorphic notion]
There are at least three different ways to axiomatize the notion of a
``category'', each with a different structure. But they are equivalent
in the sense that we can translate one axiomatization faithfully into
another. This quality (having multiple distinct axiomatizations which
are not obviously equivalent) is called \define{Cryptomorphism}.
\end{remark}

\begin{definition}[Category structure with one collection of morphisms]\mml{cat_1}
A \define{Category Structure} extends a multigraph structure $C=\structure{U,U_{1},s,t,\cdot,1}$ where
\begin{itemize}
\item the underlying set $U$ is considered to be the set of objects
  (usually written $\Ob(C)$),
\item $U_{1}$ (usually written $\Mor(C)$) is the set of morphisms,
\item $s\colon U_{1}\to U$ is the \emph{source map} 
or \emph{dom-map},
\item $t\colon U_{1}\to U$ is the \emph{target map} or
\emph{cod-map} giving us the codomain of a morphism,
\item $\cdot\colon U_{1}\times U_{1}\pto U_{1}$ is a partial function
  of morphisms called \emph{composition}, and
\item $1\colon U\to U_{1}$ is the identity morphism.
\end{itemize}
Usually we write $\Ob(C)$ for the set of objects (as a synonym for $U$
of $C$), and either $\Mor(C)$ or $\Hom(C)$ for the set of morphisms.

When we have $f\in U_{1}$ such that $s(f)=a\in U$ and $t(f)=b\in U$,
we will indicate this by saying ``$f\colon a\to b$ is a morphism of $C$''.
The set of all such morphisms is written $\hom(a,b)\subset U_{1}$. Specifically,
$\hom(a,b)=\{f\in U_{1}\mid s(f)=a\land t(f)=b\}$.

In this case, we define a \define{Category} to consist of a category
structure $\cat{C}$ such that
\begin{itemize}
\property{Category-like} for any morphism $f\in U_{1}$ and $g\in U_{1}$,
  we have $(g,f)\in\cdot$ iff $s(g)=t(f)$
\property{transitive} for any morphism $f\in U_{1}$ and $g\in U_{1}$
  such that $s(g)=t(f)$, we have $s(g\cdot f)=s(f)$ and $t(g\cdot f)=t(g)$
\property{Composition is associative} for any morphism $f\in U_{1}$ and $g\in U_{1}$
  and $h\in U_{1}$
  such that $s(h) = t(g)$ and $s(g)=t(f)$,
  we have $h\cdot(g\cdot f)=(h\cdot g)\cdot f$.
\property{reflexive} for any object $x\in U$, there exists a morphism
  from $x$ to $x$.
\property{with identities} for any object $a\in U$ there exists a
  morphism $i\in U_{1}$ such that for any object $b\in U$ we have
  (i) if there is a morphism from $a$ to $b$, then every morphism
  $g\colon a\to b$ satisfies $g\cdot i=g$, and\hfill\break
  (ii) if there is a morphism from $b$ to $a$, then every morphism
  $f\colon b\to a$ satisfies $i\cdot f=f$.
\end{itemize}
\end{definition}

\begin{remark}
The preceding definition is in the same spirit as Mac~Lane~\cite{maclane1998categories}. It may be traced back
to Grothendieck~\cite[\S4]{grothendieck1961techniques}. This has the
advantage that the definition generalizes to the notion of an
\emph{internal category} object. In fact, when we have a notion of the
category of sets, an internal category object in the category of sets
is precisely the same as the previous definition.
\end{remark}

\begin{remark}[Collections or sets]
We are lying a bit here, because we should define a category to
consist of a ``collection'' of objects and a ``collection'' of
morphisms. These colections may each separately be proper classes or sets.
But it depends on the axiomatic set theory we are working with to
describe mathematics.

This comment is also applicable to the next formalization of a category.
\end{remark}

\begin{definition}[Category structure with a family of collection of morphisms]\mml{altcat_1}
We can define a \define{Category Structure} to extend a graph
structure (\S\ref{def:graphs:graph-structure}) $C=\structure{U,A,\cdot}$
where $U$ is the underlying set (called the \emph{carrier} of $C$),
$A$ is a many-sorted set (\S\ref{def:pboole:many-sorted-set}) called
the \emph{arrows} of $C$, and
the \emph{composition} $\cdot$ is a many-sorted function (\S\ref{def:pboole:many-sorted-function}) from
$A\times A$ to $A$.

When $a\in U$ and $b\in U$, we write $\Hom(a,b)$ as another way of
writing $A_{a,b}$.

We will write $\Ob(C)$ as a synonym for the carrier $U$ of $C$.

We define a \define{Category} to consist of a category structure
$\cat{C}$ such that
\begin{itemize}
\property{transitive} for any objects $a\in U$, $b\in U$, and $c\in U$,
  if $\hom(a,b)\neq\emptyset$ and $\hom(b,c)\neq\emptyset$, then
  necessarily $\hom(a,c)\neq\emptyset$ (namely, the composition of
  $f\colon a\to b$ with $g\colon b\to c$ gives us a morphism
  $g\cdot f\colon a\to c$)
\property{associative} if the composition of $\cat{C}$ is associative,
i.e, for any objects $a\in U$ and $b\in U$ and $c\in U$ and $d\in U$, for any
morphisms $f\colon a\to b$ and $g\colon b\to c$ and $h\colon c\to d$,
we have $(h\cdot g)\cdot f=h\cdot(g\cdot f)$
\property{with units} for any objects $a\in U$ there exists a morphism
$i\in\hom(a,a)$ such that for any object $b\in U$,
 (i) for any morphism $g\colon a\to b$, we have $g\cdot i=g$, and\hfill\break
 (ii) for any morphism $f\colon b\to a$, we have $i\cdot f=f$
\end{itemize}
\end{definition}

\begin{theorem}[The two formalizations are equivalent]
For each category $\cat{C}$ with one collection of morphisms, there
exists a unique category $\cat{C}_{2}$ with a family of collections of
morphisms such that $\Ob(\cat{C})=\Ob(\cat{C}_{2})$ and
$\hom_{\cat{C}}(a,b)=\hom_{\cat{C}_{2}}(a,b)$ for all objects
$a\in\cat{C}$ and $b\in\cat{C}$ $b\in C$,  and the composition of morphisms agrees.
\end{theorem}

This is actually a \emph{metatheorem} since there is no set theoretic
function within our axiomatic set theory describing this one-to-one
correspondence.

\begin{remark}
Without loss of generality, we will simply refer to a category without
worrying about which definition to use.
\end{remark}

\begin{definition}[Common notation]
Let $\cat{C}$ be a category.

Let $x$ be any object.
We will write $x\in\cat{C}$ as an abbreviation for $x\in\Ob(\cat{C})$.

When $x\in\cat{C}$, we will write $\id_{x}\colon x\to x$ for the
identity morphism.

We will also write $\circ$ for the composition of morphisms of $\cat{C}$.

We may write $\hom_{\cat{C}}(a,b)$ or $\cat{C}(a,b)$ for the
collection of morphisms in $\cat{C}$ from $a\in\cat{C}$ to $b\in\cat{C}$.
If $\cat{C}$ is clear from context, we write $\hom(a,b)$ without fear
of ambiguity.
\end{definition}

\begin{definition}[Domain and codomain notation]\label{defn:cat:domain-and-codomain}
Let $\cat{C}$ be a category, let $f\colon a\to b$ be a morphism in
$\cat{C}$. We abuse notation and language speaking of the
\define{Domain} of $f$ as $\dom(f)=a=s(f)$ for the source of
$f$, and similarly the \define{Codomain} of $f$ is $\cod(f)=b=t(f)$ for the codomain of $f$.
\end{definition}

\begin{definition}[Large, small, and locally-small categories]
  Let $\cat{C}$ be a category.
\begin{enumerate}
\item When both the collection of objects of $\cat{C}$ forms a proper
  class and the collection of all morphisms of $\cat{C}$ forms a
  proper class, then $\cat{C}$ is called \define{Large}. If both
  collections are sets, then we call it
  \define{Small}.
\item When the collection of morphisms $\hom(a,b)$ is a set for all
  objects $a\in\cat{C}$ and $b\in\cat{C}$, then we call the category
  $\cat{C}$ \define{Locally-Small}. Note that locally-small categories
  impose no constraint on the objects being a proper class or a proper
  set. (Hence it's not hard to see that all
  small categories are also locally-small.)
\item If $\mathcal{U}$ is a Grothendieck universe, we call $\cat{C}$
  \define{$\mathcal{U}$-Small} if $\Ob(\cat{C})\in\mathcal{U}$.\mml{classes5:def 17}
\end{enumerate}
These notions depend on the particular foundations chosen for
mathematics, see Shulman~\cite{shulman2008set} for some details.

Also note that ``small objects'' (like Abelian groups, i.e.,
\emph{sets} equipped with an Abelian group structure) can be used
for $\hom(a,b)$ for the notion of ``locally small'' to still be
applicable. 
\end{definition}

\begin{example}[Some finite categories]\mml{altcat_2:def 10}
We may define the category $\cat{0}$ to consist of the empty
collection of objects and the empty collection of morphisms. This is
usually called the \define{Empty Category}.

We may define the category \mml{cat_1:def 11}$\cat{1}$ to consist of a
singleton set denoted $\{*\}$ as its collection of objects, and only
the identity morphism $\id_{*}\colon *\to *$ as its only morphism.
\end{example}

\begin{example}[Delooping a group]
\future{Let $G$ be a group. Then the delooping of $G$ is a category
  $\delooping{G}$ consisting of a single arbitrary object $*$, and
  each $g\in G$ becomes a morphism $g\colon *\to *$. The identity
  element $e\in G$ becomes the identity morphism $\id_{*}$. The
  multiplication operator of $G$ becomes the composition operator of
  $\delooping{G}$.}
\end{example}

\begin{example}[Category of sets]
The category of all sets $\Set$ has its collection of objects be the
proper class of all sets $\Ob(\Set)=\VonNeumannUniverse$, and the
morphisms are functions between sets. This is a locally-small
category, since $\hom(a,b)$ is a set in $\Set$. Note that this also
means that $\hom(a,b)\in\Ob(\Set)$.

This definition is a little ambiguous, because it depends on the exact
axiomatic set theory used. Zermelo set theory \Z/ has
$\Ob(\Set)=\VonNeumannUniverse_{2\omega}$ for example. It also depends
on whether classical logic or intuitionistic logic is used in the
axiomatization, whether the axiom of choice holds or not, whether we
demand predicative mathematics or impredicative mathematics, if the
logic is first-order or higher-order, and so on.
\end{example}

\begin{example}[Category of Ensembles]\mml{altcat_1:def 14}
Let $U$ be a nonempty set. We define the category $\Ens_{U}$ whose
objects are the elements of $U$, i.e., $\Ob(\Ens_{U})=U$. Let
$a\in\Ens_{U}$ and $b\in\Ens_{U}$, then the set of morphisms from $a$
to $b$ coincides with the set of all functions $\Ens_{U}(a,b)=\Set(a,b)$.

As discussed in Section~\ref{sec:set-theory:universes}, we could
interpret $\SETS$ as all \ZF/ sets in Tarski--Grothendieck set theory,
so we could interpret $\Ens_{\SETS}$ ``as if'' it were really $\Set$.
\end{example}

\begin{example}[Category of small categories]
\future{We may form a category $\Cat$ whose objects are all small categories
and whose morphisms are all functors between categories found in $\Cat$.}

\future{We could form a ``very large'' category $\CAT$ whose objects are all
(small or large) categories and whose morphisms are all functors
between categories found in $\CAT$. We have $\CAT\notin\CAT$ to avoid
Russell-like paradoxes.}
\end{example}

\begin{definition}[Opposite category]\mml{oppcat_1:def 1}
Let $\cat{C}$ be a category. We define the \define{Opposite} of
$\cat{C}$ to be the category $\cat{C}^{\text{op}}$ such that
\begin{enumerate}
\item $\Ob(\cat{C})=\Ob(\op{\cat{C}})$ it has the same objects as $\cat{C}$
\item The composition operator is the same in both categories $\circ_{\op{\cat{C}}}=\circ_{\cat{C}}$
\item The identity-assignment mapping is the same in both categories.
\item The source and target maps are switched, i.e.,
  $\dom_{\op{\cat{C}}}=\cod_{\cat{C}}$ and $\cod_{\op{\cat{C}}}=\dom_{\cat{C}}$.
\end{enumerate}
In particular, this means if $f\colon a\to b$ is a morphism in
$\cat{C}$, then $f\in\op{\cat{C}}(b,a)$.
\end{definition}

\begin{remark}[Duality principle: ``stuff'' and ``co-stuff'']
Any result or \emph{gadget} which may be formulated using $\cat{C}$ can also be
formulated in $\op{\cat{C}}$. We refer to this ``opposite'' version of
the gadget as a \emph{co}-gadget (i.e., we prefix the construction
with ``co-''). For example: the \emph{codomain} of a morphism is the dual to the
\emph{domain} of a morphism (\S\ref{defn:cat:domain-and-codomain}).
\end{remark}

\subsection{Functors}

\begin{definition}[Covariant functor]\mml{functor0:def 25}
Let $\cat{C}$ and $\cat{D}$ be categories.
We define a \define{(Covariant) Functor} $F\colon\cat{C}\to\cat{D}$
consists of (i) a function of objects $F_{0}\colon\Ob(\cat{C})\to\Ob(\cat{D})$ 
and (ii) a functor of morphisms $F_{1}\colon\Hom(a,b)\to\Hom(F_{0}(a),F_{0}(b))$
such that
\begin{itemize}
\property{Preserves composition} $F_{1}(g\circ f)=F_{1}(g)\circ F_{1}(f)$
when $g\colon b\to c$ and $f\colon a\to b$ are morphisms in $\cat{C}$,
\property{Preserves identity morphisms} For each object $a\in\cat{C}$,
we have $F_{1}(\id_{a})=\id_{F_{0}(a)}$.
\end{itemize}
Note that it is common to drop subscripts on functors, and simply
write $F(a)$ instead of $F_{0}(a)$, and $F(f\colon a\to b) = F_{1}(f)$.

The two properties defining a covariant functor means that commutative
diagrams are preserved by covariant functors.

When $\cat{C}=\cat{D}$, we call $F\colon\cat{C}\to\cat{C}$ an
\define{Endofunctor} of $\cat{C}$.
\end{definition}

\begin{example}[Identity functor]\mml{functor0:def 29}
Let $\cat{C}$ be a category. We have the identity functor, denoted
$\id_{\cat{C}}\colon\cat{C}\to\cat{C}$, defined on objects
$a\in\cat{C}$ by $\id_{\cat{C}}(a)=a$ and on morphisms $f\colon a\to b$
by $\id_{\cat{C}}(f)=f$.

\begin{proof}[Proof sketch]
This is a covariant functor. We will show it satisfies both properties
of a covariant functor.

Let $f\colon a\to b$ and $g\colon b\to c$ be morphisms in
$\cat{C}$. Then $\id_{\cat{C}}(g\circ f)=g\circ f=\id_{\cat{C}}(g)\circ\id_{\cat{C}}(f)$.
Hence composition is preserved.

Let $a\in\cat{C}$. Then $\id_{\cat{C}}(\id_{a})=\id_{\id_{\cat{C}}(a)}=\id_{a}$.
Hence identity morphisms are preserved.

Hence the result.
\end{proof}
\end{example}

\begin{definition}[Contravariant functor]\mml{functor0:def 27}
Let $\cat{C}$ and $\cat{D}$ be categories.
We define a \define{Contravariant Functor} $F\colon\cat{C}\to\cat{D}$
to be a covariant functor
$\cat{C}^{\text{op}}\to\cat{D}$. Specifically, this means for any
morphisms $f\colon a\to b$ and $g\colon b\to c$ in $\cat{C}$, we have
$F(g\circ f)=F(f)\circ F(g)$.
\end{definition}

\begin{example}[Hom-functors]
Let $\cat{C}$ be a category. Let $a\in\cat{C}$ be a fixed object.
Then $\hom(a,-)$ is a covariant functor with $\hom(a,f\colon b\to c)$
sending $g\colon a\to b$ to $f\circ g$.

We also have a contravariant functor $\hom(-,a)$ which sends
$b\in\cat{C}$ to $\hom(b,a)$, and sends morphisms $h\colon b\to c$ to
$\hom(h,a)\colon\hom(c,a)\to\hom(b,a)$ which are defined by sending
$g\colon c\to a$ to $\hom(h,a)(g)=g\circ h$. Observe $g\circ h\colon b\to a$.

In particular, when $\cat{C}=\Set$ we see that $\hom(-,\{0,1\})$ is
the contravariant powerset functor. How it acts on functions describes
continuous functions in topology (and measurable functions in measure
theory). This is no coincidence.
\end{example}

\begin{definition}[Faithful and Full functors]
Let $F\colon\cat{C}\to\cat{D}$ be a functor.

Then $F$ is called \mml{functor0:def 31}\define{Full} if for
all objects $x\in\cat{C}$ an $y\in\cat{C}$, we have $F\colon\cat{C}(x,y)\onto\cat{D}(F(x),F(y))$
is a surjective function.

We call $F$ \mml{functor0:def 30}\define{Faithful} if for
all objects $x\in\cat{C}$ an $y\in\cat{C}$, we have $F\colon\cat{C}(x,y)\into\cat{D}(F(x),F(y))$
is a injective function.
\end{definition}

\begin{definition}[Embedding of categories]
Let $\cat{C}$ and $\cat{D}$ be categories.
Then an \define{Embedding} of $\cat{C}$ into $\cat{D}$ is a functor
$F\colon\cat{C}\to\cat{D}$ which is both injective on objects (i.e.,
$F\colon\Ob(\cat{C})\into\Ob(\cat{D})$ is an injective function) and faithful.
\end{definition}

\begin{definition}[Composing functors]
Let $F\colon\cat{C}\to\cat{D}$ and $G\colon\cat{D}\to\cat{E}$ be functors.
We define the \define{Composition} operator of functors to give us the
composition of $F$ followed by $G$ to be the functor $G\circ F\colon\cat{C}\to\cat{E}$
such that for each object $x\in\cat{C}$ we have $(G\circ F)(x)=G(F(x))$,
and for each morphism $f\colon x\to y$ in $\cat{C}$ we have
$(G\circ F)(f\colon x\to y)=G(F(f))\colon G(F(x))\to G(F(y))$.
\end{definition}

\begin{definition}[Isomorphism and Autofunctor]\mml{functor0:def 39}
Let $\cat{C}$ and $\cat{D}$ be categories.
An \define{Isomorphism} of $\cat{C}$ into $\cat{D}$ is a functor
$F\colon\cat{C}\to\cat{D}$ such that there exists a functor
$G\colon\cat{D}\to\cat{C}$ such that $G\circ F=\id_{\cat{C}}$ and
$F\circ G=\id_{\cat{D}}$. In this case, we write $G=F^{-1}$ and refer
to $G$ as the \define{Inverse Functor} of $F$.

We call an isomorphism $F\colon\cat{C}\to\cat{C}$ an
\define{Autofunctor} or an \emph{Automorphic Functor}.
\end{definition}

\subsection{Natural Transformation}

\begin{definition}[Natural transformation]\mml{functor2:def 7}
Let $F$, $G\colon\cat{C}\to\cat{D}$ be functors.
We define a \define{Natural Transformation} $\alpha\colon F\To G$
assigns to each object $x\in\cat{C}$ a morphism $\alpha_{x}\colon F(x)\to G(x)$
in $\cat{D}$ (called the \emph{component} or \emph{coordinates} of $\alpha$ at $x$) such
that for any morphism $f\colon x\to y$ in $\cat{C}$, the following
diagram commutes in $\cat{D}$
\begin{equation}
\vcenter{\xymatrix{
F(x) \ar[r]^{F(f)}\ar[d]_{\alpha_{x}} & F(y)\ar[d]^{\alpha_{y}}\\ 
G(x) \ar[r]_{G(f)} & G(y)
}}
\end{equation}
A natural transformation is a morphism of functors.
\end{definition}

\begin{example}[Identity natural transformation]\mml{functor2:def 3}
Let $F\colon\cat{C}\to\cat{D}$ be a functor. The identity natural
transformation $\id_{F}\colon F\To F$ assigns to each object
$x\in\cat{C}$ the identity morphism $\id_{F(x)}$ in $\cat{D}$.
\end{example}

\begin{definition}[Whiskering]
Let $F,G\colon\cat{C}\to\cat{D}$ be functors, let $\alpha\colon F\To G$
be a natural transformation.
\begin{enumerate}
\item\mml{functor3:def 1} Let $H\colon\cat{D}\to\cat{E}$ be a functor. Then
\define{Whiskering} $H$ and $\alpha$ yields the natural transformation
$H\alpha\colon(H\circ F)\to(H\circ G)$ (or $H\circ\alpha$) whose
coordinates at $A$ is $H(\alpha_{A})$.
\item\mml{functor3:def 2} Let $E\colon\cat{B}\to\cat{C}$ be a functor. Then
\define{Whiskering} $\alpha$ and $E$ yields the natural transformation
$\alpha E\colon(F\circ E)\to(G\circ E)$ (or $\alpha\circ E$) whose
coordinate at $A$ is $\alpha_{F(A)}$. 
\end{enumerate}
\end{definition}

\section{Universal Properties}

\begin{definition}[Zero object]\mml{altcat_3:def 11}
Let $\cat{C}$ be a category. We define a \define{Zero Object}
$Z\in\cat{C}$ to be such that for each object $X\in\cat{C}$ there
exists a unique morphism $f\colon X\to Z$ (i.e., $Z$ is terminal) and
there exists a unique morphism $g\colon Z\to X$ (i.e., $Z$ is initial).
\end{definition}

\begin{definition}[Zero morphism]\mml{altcat_3:def 12}
Let $\cat{C}$ be a category, let $Z\in\cat{C}$ be a zero object.
Let $A$ and $B$ be objects in $\cat{C}$.
We define a \define{Zero Morphism} in $\cat{C}$ from $A$ to $B$ to be
a morphism ${0^{A}}_{B}\colon A\to B$ such that
\begin{equation}
\vcenter{\xymatrix{
    A \ar[rr]^{{0^{A}}_{B}} \ar@{.>}[rd]_{!} && B\\
    & Z\ar@{.>}[ru]_{!} &
}}
\end{equation}
commutes.
\end{definition}

\begin{definition}[Equalizer]
Let $\cat{C}$ be a category. Let $X$, $Y\in\Ob(\cat{C})$ be objects in
$\cat{C}$. Let $f$, $g\colon X\to Y$ be morphisms in $\cat{C}$. We
define the \define{Equalizer} of $f$ and $g$ to consist of an object
$E={\rm eq}(f,g)\in\cat{C}$ and a morphism $eq\colon E\to X$ such that for any
object $W$ and morphism $m\colon W\to X$ in $\cat{C}$, if $f\circ m=g\circ m$,
then there exists a unique morphism $u\colon W\to E$ such that
$eq\circ u=m$; i.e., the following diagram commutes
\begin{equation}
\vcenter{
\xymatrix{E \ar[r]^{eq} & X \ar@<0.5ex>[r]^{f} \ar@<-0.5ex>[r]_{g} & Y\\ %X\ar@<1ex>[r]^{f} \ar@<-1ex>[r]_{g} & Y\\
W\ar@{.>}[u]\ar[ur]_{m} & &}
}
\end{equation}
\end{definition}

\begin{definition}[Kernel]
Let $\cat{C}$ be a category with a zero object. Let $A$, $B$ be
objects in $\cat{C}$. Let $f\colon A\to B$ be a morphism in $\cat{C}$.
We define a \define{Kernel} of $f$ to be an object $\ker(f)\in\cat{C}$
with a map $k\colon\ker(f)\to A$ which is the equalizer of $f$ and
the zero morphism ${0^{A}}_{B}$.

These capture the intuition of a ``kernel'' that we encounter
frequently in abstract algebra (with the kernel of a group morphism,
the kernel of a ring morphism, etc.).
\end{definition}

\begin{theorem}
Let $\cat{C}$ be a category with a zero object $Z$. Let $f\colon A\to B$
be a morphism in $\cat{C}$. Then the kernel of $f$, $K = \ker(f)$, satisfies
\begin{equation}
\vcenter{\xymatrix{
A\ar[dr]^{f} & \\
K\ar[u]^{k}\ar[r]_{{0^{K}}_{B}} & B}}
\end{equation}
Moreover, it is unique (up to isomorphism) in the sense that if we had
another object $K'$ and morphism $k'\colon K'\to A$, then there exists
a unique $u\colon K'\to K$ such that
\begin{equation}
\vcenter{\xymatrix{
 & A\ar[dr]^{f} & \\
 & K\ar[u]^{k}\ar[r]_{{0^{K}}_{B}} & B\\
K'\ar@/^1.5pc/[uur]^{k'}\ar@{.>}[ur]_{u}\ar@/_1.5pc/[urr]_{{0^{K'}}_{B}} & &
}}
\end{equation}
commutes.
\end{theorem}

\begin{definition}[Coequalizer]
Let $\cat{C}$ be a category. Let $f\colon X\to Y$ and $g\colon X\to Y$
be morphisms in $\cat{C}$. We define a \define{Coequalizer} of $f$ and $g$
to consist of an object $Q$ equipped with a morphism $q\colon Y\to Q$
such that $q\circ f=q\circ g$ and for any other possible such object
$Q'$ and morphism $q'\colon Y\to Q'$ which also satisfies $q'\circ f=q'\circ g$
there exists a unique morphism $u\colon Q\to Q'$ such that $u\circ q=q'$.
That is to say, the following diagram commutes
\begin{equation}
\vcenter{
\xymatrix{
X \ar@<0.5ex>[r]^{f} \ar@<-0.5ex>[r]_{g} & Y\ar[r]^{q}\ar[dr]_{q} & Q\ar@{.>}[d]^{u}\\
                                         &                       & Q'
}
}
\end{equation}
\end{definition}

\begin{definition}[Cokernel]
Let $\cat{C}$ be a category with a zero object. Let $A$, $B$ be
objects in $\cat{C}$. Let $f\colon A\to B$ be a morphism in $\cat{C}$.
We define a \define{Cokernel} of $f$ to be an object $\coker(f)\in\cat{C}$
with a map $c\colon B\to\coker(f)$ which is the coequalizer of $f$ and
the zero morphism ${0^{A}}_{B}$.

(See Mac~Lane, Chapter~III \S3 pg~64 and Chapter~VIII \S1)
\end{definition}

\begin{example}[Cokernel of group morphism]
In $\Grp$, the cokernel of a group morphism $f\colon G\to H$ is the
quotient of $H$ by the normal closure of the image of $f$. In the case
of $\Ab$, this is just $\coker(f)=H/\im(f)$.
\end{example}


\begin{theorem}
Let $\cat{C}$ be a category with a zero object. Let $f\colon A\to B$
be a morphism in $\cat{C}$. A cokernel of $f$ in $\cat{C}$ is an
object $Q$ and a morphism $q\colon B\to Q$ such that for any other
object $Q'$ and morphism $q'\colon B\to Q'$ there exists a unique
$u\colon Q\to Q'$ making the following diagram commute:
\begin{equation}
\vcenter{
\xymatrix{
             & B\ar[d]^{q}\ar@/^1.5pc/[ddr]^{q'} & \\
A\ar[ur]^{f}\ar@/_1.5pc/[drr]_{{0^{A}}_{Q'}}\ar[r]_{{0^{A}}_{Q}} & Q \ar@{.>}[dr]_{u}      &\\
            &                          & Q'}
}
\end{equation}
\end{theorem}

\section{Abelian Categories}

\begin{remark}[Abelian categories is where we can do homology]
``Abelian'' categories form the most general setting where we can do
homological algebra. That's why we're interested in them!
\end{remark}

\begin{definition}[Preadditive categories]
We call a category $\cat{C}$ \define{Preadditive} iff
\begin{itemize}
\property{$\Ab$-enriched:} for all objects
$A\in\cat{C}$ and $B\in\cat{C}$, $\hom_{\cat{C}}(A,B)$ has the
structure of an Abelian group; and
\property{Composition is bilinear} for all morphisms $f\colon B\to C$ and $f'\colon B\to C$
in $\cat{C}$ and for all $g\colon A\to B$ in $\cat{C}$, we have
$(f+f')\circ g=(f\circ g)+(f'\circ g)$
and for all $h\colon C\to D$ we have $h\circ(f+f')=(h\circ f)+(h\circ f')$.
\end{itemize}
\end{definition}

\begin{example}
The category $\Ab$ is pre-additive.
\end{example}

\begin{definition}[Additive categories]
A category $\cat{C}$ is called \define{Additive} iff all the following holds
\begin{itemize}
\property{Pre-additive} $\cat{C}$ is pre-additive, and
\property{Zero-objects} $\cat{C}$ has zero objects, and
\property{Finite products} $\cat{C}$ has finite products, and
\property{Finite coproducts} $\cat{C}$ has finite coproducts.
\end{itemize}
Observe that pre-additive categories with zero objects implies every
hom-set is nonempty.
\end{definition}

\begin{example}
The category $\Ab$ of Abelian groups is an additive category. It's
pre-additive. The trivial group is a zero object. Also products
coincide with coproducts in $\Ab$.
\end{example}

\begin{definition}[Abelian categories]
Mac~Lane's \emph{Categories for the Working Mathematician} (Ch.~VIII,
\S3, pg.198) offers the following definition:

An \define{Abelian} category $\cat{A}$ is an Ab-category satisfying
the following conditions:
\begin{enumerate}
\item $\cat{A}$ has a null object
\item $\cat{A}$ has binary biproducts
\item Every arrow in $\cat{A}$ has a kernel and cokernel
\item Every monid arrow is a kernel, and every epic arrow is a cokernel
\end{enumerate}
Equivalently, we can demand the following:
\begin{itemize}
\property{Additive} $\cat{A}$ is an additive category
\property{Kernel} Every morphism $f\colon A\to B$ in $\cat{A}$ has a kernel
\property{Cokernel} Every morphism $f\colon A\to B$ in $\cat{A}$ has a cokernel
\property{Parallel morphisms are isomorphisms} For any morphism
$f\colon A\to B$, the morphism $\phi\colon\coim(f)\to\im(f)$ parallel
to $f$ is an isomorphism.
\end{itemize}
The first three properties define ``pre-Abelian'' categories.
\end{definition}

